Research Notes

### 2018-06-17 writups/notes.tex
check in partial notes

@@

\section{Chebyshev Polynomials}

>>>>

\section{Volumetric Forcing}

Due to the structure of \lstinline{Dedalus}, we must use Chebyshev polynomials
in the vertical direction (to be able to enforce BCs). But as discussed in
\autoref{s:chebyshev}, the grid spacing of the Chebyshev polynomials near
boundaries goes as $1/N^2$, which means that interface forcing necessitates an
extremely strict Courant condition to capture evolution near the interface; this
is the explanation for the divergences we were seeing in earlier sections.

Instead, we employ volumetric forcing. As a toy problem of seeing how this
works, we will study the restricted 1D wave equation.

\subsection{1D Wave Equation}

Consider the forced 1D wave equation (we set $c = 1$)
\begin{equation}
    \ptd{u(x, t)}{t} - \ptd{u(x, t)}{x} = f(x, t).
\end{equation}
Note that the only allowed propagation modes here are $u \sim e^{iC\p*{x \pm
t}}$ for some constant $C$.

We wish to choose $f(x, t)$ such that $u\p*{x \to -\infty} = 0$ while $u\p*{x
\to +\infty} \sim e^{i\p*{x - t}}$. It is clear that if we consider the Fourier
transform $\tilde{f}(k, t)$, the response in $\tilde{u}(k, t)$ will also contain
components for each of the nonzero $\tilde{f}(k, t)$ components. Thus, if we
want a pure $e^{i(x - t)}$ outgoing wave, we should have a pure $\tilde{f}(k, t)
\propto \delta\p*{k - 1}$. However, it is clear that driving the oscillator
exactly on resonance produces divergent oscillation amplitudes, and moreover it
is impossible to drive with an $f(x, t)$ that doesn't vanish as $x \to \pm
\infty$.

It is nonetheless clear that $\tilde{f}(k, t)$ should be strongly peaked at $k =
1$ and small elsewhere. We can thus consider a form $f(x, t) \propto
e^{-x^2/2\sigma^2} e^{i(x - t)}$ such that $\sigma \gg 1$, i.e.\ the envelope
contains many wavelengths. The Fourier transform $\tilde{f}(k, t)$ is then a
Gaussian with width $1/\sigma \ll 1$ centered at $k = 1$ (convolution).

More precisely, consider $f(x, t) = f_0e^{-x^2/2\sigma^2} e^{i(x - t)}$ and $u(x,
t) = u_0(x)e^{i(x - t)}$ such that $u_0(x \to \infty) = 0$. Substituting this
into the inhomogeneous PDE gives
\begin{equation}
    i\pd{u_0(x)}{x} - \ptd{u_0(x)}{x} = f_0e^{-x^2/2\sigma^2}.
\end{equation}
We note that if $\sigma \gg 1$ then roughly $\pd{}{x} \ll 1$ or
\begin{equation}
    \pd{u_0(x)}{x} \approx -if_0e^{-x^2/2\sigma^2}.
\end{equation}
Enforcing $u_0(x \to -\infty) = 0$, then $\abs*{u_0(x \to +\infty)} =
f_0\sigma\sqrt{2\pi}$.

@@

\section{Chebyshev Polynomials}\label{s:chebyshev}
